---
title: "climcropr soil"
author: "Andy South"
date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
#        fig_caption: yes
output: pdf_document        
vignette: >
  %\VignetteIndexEntry{climcropr intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=7)

# library(tidyverse)
# library(forcats)
# library(sp) #for maps
# library(tmap) #for maps
library(climcropr)
library(raster)
library(rnaturalearth) #for maps
library(dismo) #for ecocrop
library(knitr) #for tables with kable

# extent object for use in plots
ext <- raster::extent(-180,180,-60,85)

#BEWARE names seem to be different in new df_ecocrop$COMNAME
#e.g. broom-corn is sorghum
test_crops <- c('potato','broom-corn','maize','oat','barley','apple','lentil','groundnut','sugarbeet', "rice","chickpea")

#test_crops <- c('potato','broom-corn','maize','barley')

```

## looking at soil ph and other limits from ecocrop & applying to world harmonised soil database
## other soil parameters on end of document

## ecocrop has drainage, depth and texture. Each seems to have a required (R) and optimal value.
## They don't quite fit with data available in the harmonised world soil database and not clear to me whether they will be useful for us.

### min & max pH for selection of crops from new scraped ecocrop data base
```{r, echo=FALSE, warnings=FALSE, message=FALSE, eval=TRUE}

df_crop_ph <- data.frame(cropname=test_crops)

for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]
  df_crop_ph$phmin[cropnum] <- get_ecocrop(cropname,'phmin')
  df_crop_ph$phmax[cropnum] <- get_ecocrop(cropname,'phmax')
  df_crop_ph$phopmn[cropnum] <- get_ecocrop(cropname,'phopmn')
  df_crop_ph$phopmx[cropnum] <- get_ecocrop(cropname,'phopmx')  
  #cat(cropname, " min,max = ", df_crop_ph$phmin[cropnum], ",", df_crop_ph$phmax[cropnum])
}

kable(df_crop_ph)

```

### crop pH suitability maps, 1=suitable, 2=optimal
```{r, echo=FALSE, warnings=FALSE, message=FALSE, eval=FALSE}
#TODO re-enable eval=TRUE, disabling now to look at non pH params

#test_crops <- c('potato','broom-corn','maize','barley')


for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]


  rst <- suit_soil_ph(cropname, plot=FALSE)
  
  par(mar=c(0,0,2,0)) #bltr
  plot(rst, main=paste0(cropname," soil pH suit=1, opt=2"), ext=ext)
  plot(ne_countries(), add=TRUE, border='grey', lwd=0.1)
  
}

```

\pagebreak

## crop drainage requirements from ecocrop 
### don't seem that they will be very useful for discrimintaing areas

```{r, echo=FALSE, warnings=FALSE, message=FALSE}

dfvals <- data.frame(values=levels(df_ecocrop$DRA), stringsAsFactors = FALSE)
kable(dfvals)

df_crop_drain <- data.frame(cropname=test_crops)

for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]
  df_crop_drain$DRA[cropnum] <- get_ecocrop(cropname,'DRA')
  df_crop_drain$DRAR[cropnum] <- get_ecocrop(cropname,'DRAR')
  #cat(cropname, " min,max = ", df_crop_ph$phmin[cropnum], ",", df_crop_ph$phmax[cropnum])
}

#for example crops
kable(df_crop_drain)

```


The harmonised world soil database has drainage data in these categories which don't seem to relate directly.

1. 150 mm water per m of the soil unit
2. 125 mm
3. 100 mm 
4. 75 mm 
5. 50 mm 
6. 15 mm 
7. 0 mm


\pagebreak

## soil depth requirements from ecocrop 
### don't seem that they will be very useful for discrimintaing areas

```{r, echo=FALSE, warnings=FALSE, message=FALSE}

dfvals <- data.frame(values=levels(df_ecocrop$DEP), stringsAsFactors = FALSE)
kable(dfvals)

df_crop_depth <- data.frame(cropname=test_crops)

for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]
  df_crop_depth$DEP[cropnum] <- get_ecocrop(cropname,'DEP')
  df_crop_depth$DEPR[cropnum] <- get_ecocrop(cropname,'DEPR')
  #cat(cropname, " min,max = ", df_crop_ph$phmin[cropnum], ",", df_crop_ph$phmax[cropnum])
}

#for example crops
kable(df_crop_depth)

```

World harmonised soil database has :

REF_DEPTH	   
Reference soil depth of all soil units are set at 100 cm, except for Rendzinas and Rankers of FAO-74 and Leptosols of FAO-90, where the reference soil depth is set at 30 cm, and for Lithosols of FAO-74 and Lithic Leptosols of FAO-90, where it is set at 10 cm.

ROOTS	    
Depth of obstacles to roots.	

0.  no information, 
1.  no obstacles to roots between 0 and 80 cm depth, 
2.  obstacles to roots between 60 and 80 cm depth, 
3.  obstacles between 40 and 60 cm, 
4.  20 and 40 cm, 
5.  0 and 80 cm, 
6.  0 and 20 cm.



\pagebreak

## soil texture requirements from ecocrop 

```{r, echo=FALSE, warnings=FALSE, message=FALSE}

dfvals <- data.frame(values=levels(df_ecocrop$TEXT), stringsAsFactors = FALSE)
kable(dfvals)

df_crop_texture <- data.frame(cropname=test_crops)

for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]
  df_crop_texture$TEXT[cropnum] <- get_ecocrop(cropname,'TEXT')
  df_crop_texture$TEXTR[cropnum] <- get_ecocrop(cropname,'TEXTR')
  #cat(cropname, " min,max = ", df_crop_ph$phmin[cropnum], ",", df_crop_ph$phmax[cropnum])
}

#for example crops
kable(df_crop_texture)

```

### harmonised world soil database has separate layers for % weight of :

1. clay
2. gravel
3. sand
4. silt 
5. organic carbon

Not clear how this could be related to ecocrop texture.
