---
title: "climcropr intro"
author: "Andy South"
date: "`r Sys.Date()`"
#output: rmarkdown::html_vignette
#        fig_caption: yes
output: pdf_document        
vignette: >
  %\VignetteIndexEntry{climcropr intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=7)

# library(tidyverse)
# library(forcats)
# library(sp) #for maps
# library(tmap) #for maps
# library(knitr) #for tables with kable
library(climcropr)
library(raster)
library(rnaturalearth) #for maps
library(dismo) #for ecocrop

# extent object for use in plots
ext <- extent(-180,180,-60,80)

```

This is a data driven document with hidden code that generates the figures. 

To create an ecocrop maximum suitability map for a crop you can use the `ecocrop_a_raster` function.

This accepts raster stacks of climate data for monthly inputs.

Currently the options are :

1. st_clim_all : a raster stack with 36 layers named on01–on12 for min, oa01-oa12 for mean and op01-012 for precip.
1. st_tmin, st_tavg, st_prec : 3 separate raster stacks of 12 layers each, names don't matter (order does)


```{r, eval=FALSE, echo=TRUE}

# call function using example climate data in the package
data("st_clim")
ec_out_sorghum <- ecocrop_a_raster('Sorghum (high altitude)', 
                                   st_clim_all = st_clim, 
                                   rainfed = TRUE, 
                                   filename = "test_out",
                                   overwrite = TRUE )

#separating out tmin, avg, precip etc. from the example input file
#this is also done within the ecocrop_a_raster() function
#I've put it here too allow experimenting with replacing tavg with tmax
    month_nums <- c(paste0('0',1:9),10:12)
    names_tmin <- paste0('on',month_nums)
    names_tmax <- paste0('ox',month_nums)
    names_tavg <- paste0('oa',month_nums)
    names_prec <- paste0('op',month_nums)

    st_tmin <- raster::subset(st_clim_all, subset=names_tmin)
    st_tmax <- raster::subset(st_clim_all, subset=names_tmax)
    st_tavg <- raster::subset(st_clim_all, subset=names_tavg)
    st_prec <- raster::subset(st_clim_all, subset=names_prec)

# call function using individual raster stacks for monthly tmin, tavg and precip
# in this case experimenting with using tmax instead of tavg
ec_out_potato <- ecocrop_a_raster('potato', 
                                   st_tmin = st_tmin, 
                                   st_tavg = st_tmax, #BEWARE tavg set from tmax
                                   st_prec = st_prec, 
                                   rainfed = TRUE, 
                                   filename = "test_out2",
                                   overwrite = TRUE )
```



```{r, eval=FALSE, echo=FALSE, warnings=FALSE, message=FALSE, include=FALSE}
# val=FALSE above because models take a few minutes to run
# run the models

#separating out tmin, avg, precip etc. from input file
#this is also done within the ecocrop_a_raster() function
#I've put it here too allow experimenting with replacing tavg with tmax
    month_nums <- c(paste0('0',1:9),10:12)
    names_tmin <- paste0('on',month_nums)
    names_tmax <- paste0('ox',month_nums)
    names_tavg <- paste0('oa',month_nums)
    names_prec <- paste0('op',month_nums)

    st_tmin <- raster::subset(st_clim_all, subset=names_tmin)
    st_tmax <- raster::subset(st_clim_all, subset=names_tmax)
    st_tavg <- raster::subset(st_clim_all, subset=names_tavg)
    st_prec <- raster::subset(st_clim_all, subset=names_prec)

#sorghum using tavg
ec_out_sorghum <- ecocrop_a_raster('Sorghum (high altitude)', 
                                   st_clim_all = st_clim, 
                                rainfed = TRUE, 
                                filename = "inst\\extdata\\ec_out_sorghum",
                                overwrite = TRUE )

#sorghum using tmax
ec_out_sorghum_max <- ecocrop_a_raster('Sorghum (high altitude)', 
                                   st_tmin = st_tmin, 
                                   st_tavg = st_tmax, #BEWARE using tmax for tavg
                                   #st_tavg = st_tavg, 
                                   st_prec = st_prec, 
                                   rainfed = TRUE, 
                                   filename = "inst\\extdata\\ec_out_sorghum_max",
                                   overwrite = TRUE )

ec_out_potato <- ecocrop_a_raster('potato', 
                                   st_clim_all = st_clim, 
                                   rainfed = TRUE, 
                                   filename = "inst\\extdata\\ec_out_potato_rain",
                                   overwrite = TRUE )

#potato using tmax
ec_out_potato_max <- ecocrop_a_raster('potato', 
                                       st_tmin = st_tmin, 
                                       st_tavg = st_tmax, #BEWARE using tmax for tavg
                                       #st_tavg = st_tavg, 
                                       st_prec = st_prec, 
                                       rainfed = TRUE, 
                                       filename = "inst\\extdata\\ec_out_potato_max",
                                       overwrite = TRUE )

#sugar beet
ec_out_beet <- ecocrop_a_raster('sugar beet', 
                                  st_clim_all = st_clim, 
                                  rainfed = TRUE, 
                                  filename = "inst\\extdata\\ec_out_beet",
                                  overwrite = TRUE )

#sugar beet using tmax
ec_out_beet_max <- ecocrop_a_raster('sugar beet', 
                                      st_tmin = st_tmin, 
                                      st_tavg = st_tmax, #BEWARE using tmax for tavg
                                      #st_tavg = st_tavg, 
                                      st_prec = st_prec, 
                                      rainfed = TRUE, 
                                      filename = "inst\\extdata\\ec_out_beet_max",
                                      overwrite = TRUE )

ec_out_maize <- ecocrop_a_raster('maize', 
                                   st_clim_all = st_clim, 
                                   rainfed = TRUE, 
                                   filename = "inst\\extdata\\ec_out_maize",
                                   overwrite = TRUE )

ec_out_barley <- ecocrop_a_raster('barley', 
                                   st_clim_all = st_clim, 
                                   rainfed = TRUE, 
                                   filename = "inst\\extdata\\ec_out_barley",
                                   overwrite = TRUE )
#irrigated
ec_out_potato_irig <- ecocrop_a_raster('potato',
                                   st_clim_all = st_clim,
                                   rainfed = FALSE,
                                   filename = "inst\\extdata\\ec_out_potato_irig",
                                   overwrite = TRUE )
ec_out_maize_irig <- ecocrop_a_raster('maize',
                                   st_clim_all = st_clim,
                                   rainfed = FALSE,
                                   filename = "inst\\extdata\\ec_out_maize_irig",
                                   overwrite = TRUE )

```

# Initial experimenting with outputs

### SORGHUM using tmean
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}
#```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=8.5, fig.width=6}

#read in files output by ecocrop
ec_out_sorghum <- raster(system.file("extdata", "ec_out_sorghum.grd" , package = "climcropr"))
ec_out_sorghum_max <- raster(system.file("extdata", "ec_out_sorghum_max.grd", package = "climcropr"))

par(mar=c(0,0,0,0))
plot(ec_out_sorghum, horizontal=TRUE, axes=FALSE, legend=TRUE, ext=ext)

```

### SORGHUM using tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}

par(mar=c(0,0,0,0))
plot(ec_out_sorghum_max, horizontal=TRUE, axes=FALSE, ext=ext)

```

### SORGHUM difference between using tmean and tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#mean - max
par(mar=c(0,0,0,0))
plot(ec_out_sorghum - ec_out_sorghum_max, horizontal=TRUE, axes=FALSE, ext=ext)

```


### SORGHUM mean - max for Africa
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

par(mar=c(0,0,0,0))
#set region
#plot(ne_countries(continent=c('Africa','Asia')))
plot(ne_countries(continent=c('Africa')), axes=FALSE)
plot(ec_out_sorghum - ec_out_sorghum_max, add=TRUE, horizontal=TRUE, axes=FALSE)
#add countries on top of raster
plot(ne_countries(continent='Africa'), add=TRUE, border='grey', lwd=0.5)

```

### potato using tmean
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}
#```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=8.5, fig.width=6}

#read in files output by ecocrop
ec_out_potato <- raster(system.file("extdata", "ec_out_potato.grd", package = "climcropr"))
ec_out_potato_max <- raster(system.file("extdata", "ec_out_potato_max.grd", package = "climcropr"))

par(mar=c(0,0,0,0))
plot(ec_out_potato, horizontal=TRUE, axes=FALSE, legend=TRUE, ext=ext)

```

### potato using tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}

par(mar=c(0,0,0,0))
plot(ec_out_potato_max, horizontal=TRUE, axes=FALSE, ext=ext)

```

### potato difference between using tmean and tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#mean - max
par(mar=c(0,0,0,0))
plot(ec_out_potato - ec_out_potato_max, horizontal=TRUE, axes=FALSE, ext=ext)

```


### potato mean - max for Africa
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

par(mar=c(0,0,0,0))
#set region
#plot(ne_countries(continent=c('Africa','Asia')))
plot(ne_countries(continent=c('Africa')), axes=FALSE)
plot(ec_out_potato - ec_out_potato_max, add=TRUE, horizontal=TRUE, axes=FALSE)
#add countries on top of raster
plot(ne_countries(continent='Africa'), add=TRUE, border='grey', lwd=0.5)

```

### beet using tmean
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}
#```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=8.5, fig.width=6}

#read in files output by ecocrop
ec_out_beet <- raster(system.file("extdata", "ec_out_beet.grd", package = "climcropr"))
ec_out_beet_max <- raster(system.file("extdata", "ec_out_beet_max.grd", package = "climcropr"))

par(mar=c(0,0,0,0))
plot(ec_out_beet, horizontal=TRUE, axes=FALSE, legend=TRUE, ext=ext)

```

### beet using tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE, fig.height=3.5}

par(mar=c(0,0,0,0))
plot(ec_out_beet_max, horizontal=TRUE, axes=FALSE, ext=ext)

```

### beet difference between using tmean and tmax
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#mean - max
par(mar=c(0,0,0,0))
plot(ec_out_beet - ec_out_beet_max, horizontal=TRUE, axes=FALSE, ext=ext)

```


### beet mean - max for Africa
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

par(mar=c(0,0,0,0))
#set region
#plot(ne_countries(continent=c('Africa','Asia')))
plot(ne_countries(continent=c('Africa')), axes=FALSE)
plot(ec_out_beet - ec_out_beet_max, add=TRUE, horizontal=TRUE, axes=FALSE)
#add countries on top of raster
plot(ne_countries(continent='Africa'), add=TRUE, border='grey', lwd=0.5)

```


### min & max pH for selection of crops 
### from new scraped ecocrop data base
```{r, echo=FALSE, warnings=FALSE, message=FALSE}

#BEWARE names seem to be different in new df_ecocrop$COMNAME
#e.g. broom-corn is sorghum
test_crops <- c('potato','broom-corn','maize','oat','barley','apple','lentil','groundnut','sugarbeet', "rice","chickpea")

#test_crops <- c('potato','broom-corn','maize','barley')

df_crop_ph <- data.frame(cropname=test_crops,
                         phmin=NA, phmax=NA,
                         phopmin=NA, phopmax=NA)

for(cropnum in 1:length(test_crops))
{
  cropname <- test_crops[cropnum]
  df_crop_ph$phmin[cropnum] <- get_phmin(cropname)
  df_crop_ph$phmax[cropnum] <- get_phmax(cropname) 
  df_crop_ph$phopmin[cropnum] <- get_phopmin(cropname)
  df_crop_ph$phopmax[cropnum] <- get_phopmax(cropname)   
  cat(cropname, " min,max = ", df_crop_ph$phmin[cropnum], ",", df_crop_ph$phmax[cropnum])
}




```

